Robustness of MinD oscillation in \emph{Escherichia coli} with diverse cell shapes

### 2014-09-11 paper/paper.tex
edits to mannik section of paper and small print statement added to dens-corr Tested everything on quipu.

We have made and analyzed movies out of color plots of MinD
concentration that are constructed as those in Figure~\ref{image-p},
integrated along the flattened axis, in which each frame is seperated
in simulation time by a half second. Figure~\ref{randst-plot-ave}
illustrates what is seen in these movies. In his work Mannik presents
arrow plots in which the arrow heads show the location of sequential
MinD maxima within the cell. His plots for his two abnormal cell
shapes are shown in the upper 'experiment' section of
Figure~\ref{randst-plot-ave}.  Following this, we have created plots
that show maxima that are global in space and local in time.  Of these
maxima we only show those which have greater concentration than a
chosen cutoff concentration, and also ignore maxima that appear
temporalily and spatially adjacent to preceding maxima, in order to
stop single spikes from appearing as many adjecent ones.  The overal
effect is to accurately describe what is seen in our movies.

We have found that the deterministic simulations of the protein
reaction sequence discussed above results in robust oscillatory
behavoir of polar selection and oscillation in not only symmetrical
shapes, such as the wildtype pill shape, but also in very
assymetrical, flatttened cells, such as those studied by Mannik.
After the lower stability limit of
section~\ref{sec:stability-analysis} \fixme{this needs to be covered}
has been surpassed, this qualitative behavoir doesn't seem to have
size dependence until cells reach sizes that have more than twice the
dimensions of the Mannik cells.  At this point the robust
oscillations, while still evident, begin to lose their consistanty.
The Mannik reported cells already seem to be rather large (wild type
cells are typically smaller than those observed my Mannik
\cite{kubitschek1990cell,kubitschek1968linear,mannik2012robustness}),
so we left off simulating cells any larger than twice this size.

The movies show sharp peaks of MinD in specific locatations, typically at
the edges of the cell and in regions of high radius of curvature,
followed by a spreading out of the MinD out into the cell, with no
real bias towards any location, until it aggregates again in a
specific location at the opposite edge of the cell, maximizing there.
The pattern is repeated again and again with surprising consistency,
which can be seen on the right of Figure~\ref{box-mannik}.  This is
the same plotting method as our box plots above, here showing
stochastic and deterministic simulations of cells that are the same
shape and size of one of Mannik's reported cells.

The results of these simulations differ significantly with Mannik's
experimental results that show much less regular behavoir, with a
large variation in the location of maxima.

>>>>

We have made and analyzed movies out of color plots of MinD
concentration that are constructed as those in Figure~\ref{image-p},
integrated along the flattened axis, in which each frame is seperated
in simulation time by a half second. Figure~\ref{randst-plot-ave}
illustrates what is seen in these movies. In his work Mannik presents
arrow plots in which the arrow heads show the location of sequential
MinD maxima within the cell. His plots for his two abnormal cell
shapes are shown in the upper 'experiment' section of
Figure~\ref{randst-plot-ave}.  Following this, we have created plots
that show maxima that are global in space and local in time.  Of these
maxima we only show those which have greater concentration than a
chosen cutoff concentration, and also ignore maxima that appear
temporaly and spatially adjacent to preceding maxima, in order to
stop single spikes from appearing as many adjecent ones.  The overall
effect is to accurately describe what is seen in our movies.

We have found that the deterministic simulations of the protein
reaction sequence discussed above results in robust oscillatory
behavoir of polar selection and oscillation in not only symmetrical
shapes, such as the wildtype pill shape, but also in very
assymetrical, flatttened cells, such as those studied by Mannik.
After the lower size stability limit of
section~\ref{sec:stability-analysis} \fixme{this needs to be covered}
has been surpassed, this qualitative behavoir doesn't seem to have
size dependence until cells reach sizes that have more than twice the
dimensions of the Mannik cells.  At this point the robust
oscillations, while still evident, begin to lose their consistency.
The Mannik reported cells already seem to be rather large (wild type
cells are typically smaller than those observed my Mannik
\cite{kubitschek1990cell,kubitschek1968linear,mannik2012robustness}),
so we left off simulating cells any larger than twice this size.

Looking at Mannik's shape A in Figure~\ref{randst-plot-ave}, we can
see that the deterministic simulation selects out two poles, one at
either end of the cell, and that the MinD does not ever maximize in
any other location.  All of the flattened cells shown in the figure
span a time of 350s (as do Mannik's plots), which corresponds to
\fixme{number of periods} periods for shape A.  The movies show sharp
peaks at these poles, followed by a spreading out of the MinD out into
the cell, with no real bias towards any location, until it aggregates
again at the opposite pole, maximizing there.  The pattern is repeated
again and again with surprising consistency, \fixme{talk about the
  time correlation for the deterministic, once the plots are added.}

The results of these deterministic simulations for shape A differ
significantly with Mannik's experimental results that show much less
regular behavoir, with a large variation in the location of maxima,
shown in the experiment section of Figure~\ref{randst-plot-ave}.

The deterministic results of shape B are slightly different, as can be
seen in the figure.  Here we can see that the proteins maximize on the
rightward cell wall while on the way to their corner maxima locations.
Watching the movies, the proteins seem to travel along the wall, from
corner to corner and back.  This behavoir is still more irregular than
that of the Mannik experiments, which seem to show some right-left
back and forth behavoir in a few areas, as opposed to the straight
pole to pole (traveling along the wall).

These comparisons illustrate our first finding, which is that in
irregular cell shapes, the deterministic method of simulation is indeed
inadequate in representing and predicting experimental behavoir.
